Kids PM Summary

### 2019-12-12 pm_summary.Rmd
pm cleaning

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

@@

# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Comments" = first(qaqccomments))

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(area, dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

>>>>

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(lubridate)
jvPalette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
            "Act vs Exp" = first(actual_vs_expected),
            "Comments" = first(qaqccomments))
write_csv(summary_id, "output/kids_summary_id.csv")

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(area, dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))

histogram_id <- kids_pm %>% 
  filter(home_winter_id == "10" & area == "AK") %>% 
  ggplot() + 
    geom_histogram(aes(pm), bins = 50) +
    theme_classic()
histogram_id

##### 2 Functions to look at PM data across time #####

# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jvPalette)

gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

pm_time_plot
}

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "228", location = "NN")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check <- kids_pm %>% 
  filter(area == "NN" & home_winter_id == "224") %>% 
  filter(pm < 0)
```
